Light and growth form interact to shape stomatal ratio among British angiosperms

### 2017-10-04 ms/ms.Rnw
updated light-growth form results

I did not use the raw covariance, but rather estimated the phylogenetic covariance matrix between \el, $\mathrm{sd_{ab}}$, and $\mathrm{sd_{ad}}$ using a multivariate Ornstein-Uhlenbeck model fit in \pkg{Rphylopars} version 0.2.9 \citep{Goolsby_etal_2016, Goolsby_etal_2017}. From the covariance matrix, I estimated the contribution of abaxial density, adaxial density, and their covariance as:

@@

I also wanted to test whether light-mediated evolution of stomatal ratio acted mostly by 1) increasing adaxial stomatal density while maintaining abaxial density, or 2) keeping total stomatal density the same, but shifting a greater proportion to the adaxial surface. The first scenario predicts that the phylogenetic regression of \el~ on $\mathrm{sd_{ad}}$ is stronger than that for $\mathrm{sd_{ab}}$. The second scenario predicts that \el~acts similarly on both and that there is a negative covariance ($\mathrm{Cov(sd_{ad}, sd_{ab}) < 0}$). I tested these competing predictions by fitting a very simple phylogenetic SEM \citep[see][for a similar approach]{Mason_etal_2016}. The model uses the phylogenetic covariance matrix, as described above, to simultaneously estimate regressions of \el~on $\mathrm{sd_{ad}}$ and $\mathrm{sd_{ab}}$ while allowing covariance between them (i.e. estimating $\mathrm{Cov(sd_{ad}, sd_{ab})}$). To fit the SEM, I used the R package \pkg{lavaan} version 0.5-23.1097 \citep{Rosseel_2012}. I tested whether parameter estimates were significantly different from zero using $z$-scores.

@@

\subsection*{Light tolerance varies among Raunki\ae r life forms}

Ellenberg light indicator values (\el) differed significantly among life forms (Fig.~\ref{fig:lf-light};ANOVA - $F_{\Sexpr{aovLFvEL$Df[1]}, \Sexpr{aovLFvEL$Df[2]}}$ = \Sexpr{round(aovLFvEL$`F value`[1], 1)}, $P$ = \Sexpr{pval2latex(aovLFvEL$`Pr(>F)`[1])}). Therophytes (annuals), hemicryptophytes (perennial herbs with buds near the soil surface), and chamaephytes (subshrubs) had greater \els~than phanerophytes (large woody plants) and geophytes (perennial herbs with storage organs) (Fig.~\ref{fig:lf-light}).

>>>>

I did not use the raw covariance, but rather estimated the phylogenetic covariance matrix between \el, $\mathrm{sd_{ab}}$, and $\mathrm{sd_{ad}}$ using a multivariate Ornstein-Uhlenbeck model fit in \pkg{Rphylopars} version 0.2.9 \citep{Goolsby_etal_2016, Goolsby_etal_2017}. The phylogenetic covariance measures how strongly a set of traits evolve together over macroevolutionary timescales. From the covariance matrix, I estimated the contribution of abaxial density, adaxial density, and their covariance as:

@@

I also wanted to test whether light-mediated evolution of stomatal ratio acted mostly by 1) increasing adaxial stomatal density while maintaining abaxial density, or 2) keeping total stomatal density the same, but shifting a greater proportion to the adaxial surface. The first scenario predicts that the phylogenetic regression of \el~ on $\mathrm{sd_{ad}}$ is stronger than that for $\mathrm{sd_{ab}}$. The second scenario predicts that \el~acts similarly on both and that there is a negative covariance ($\mathrm{Cov(sd_{ad}, sd_{ab}) < 0}$). I tested these competing predictions by fitting a very simple phylogenetic SEM \citep[see][for a similar approach]{Mason_etal_2016}. In general, SEMs attempt to determine whether variables are related causally or whether a relationship is mediated by another correlated variable. Phylogenetic SEMs use the phylogenetic covariance matrix, as described above, rather than the raw covariance. Here, I used a phylogenetic SEM to simultaneously estimate regressions of \el~on $\mathrm{sd_{ad}}$ and $\mathrm{sd_{ab}}$ while allowing covariance between them (i.e. estimating $\mathrm{Cov(sd_{ad}, sd_{ab})}$). To fit the SEM, I used the R package \pkg{lavaan} version 0.5-23.1097 \citep{Rosseel_2012}. I tested whether parameter estimates were significantly different from zero using $z$-scores.

@@

\subsection*{Light tolerance varies among growth forms}

Ellenberg light indicator values (\el) differed significantly among growth forms. Among Raunki\ae r life forms, therophytes (annuals), hemicryptophytes (perennial herbs with buds near the soil surface), and chamaephytes (subshrubs) had greater \els~than phanerophytes (woody plants) and geophytes (perennial herbs with storage organs) (Fig.~\ref{fig:lf-light}; ANOVA - $F_{\Sexpr{aovLFvEL$Df[1]}, \Sexpr{aovLFvEL$Df[2]}}$ = \Sexpr{round(aovLFvEL$`F value`[1], 1)}, $P$ = \Sexpr{pval2latex(aovLFvEL$`Pr(>F)`[1])}). Likewise, herbaceous plants (annual, biennial, and perennials) had greater \els~than shrubs and trees (Fig.~\ref{fig:gf-light}; ANOVA - $F_{\Sexpr{aovGFvEL$Df[1]}, \Sexpr{aovGFvEL$Df[2]}}$ = \Sexpr{round(aovGFvEL$`F value`[1], 1)}, $P$ = \Sexpr{pval2latex(aovGFvEL$`Pr(>F)`[1])})

### 2017-10-04 ms/ms.tex
updated light-growth form results

\subsection*{Light tolerance varies among Raunki\ae r life forms}

Ellenberg light indicator values (\el) differed significantly among life forms (Fig.~\ref{fig:lf-light};ANOVA - $F_{4, 367}$ = 18.3, $P$ = 1.1 $\times10^{-13}$). Therophytes (annuals), hemicryptophytes (perennial herbs with buds near the soil surface), and chamaephytes (subshrubs) had greater \els~than phanerophytes (large woody plants) and geophytes (perennial herbs with storage organs) (Fig.~\ref{fig:lf-light}).

>>>>

\subsection*{Light tolerance varies among growth forms}

Ellenberg light indicator values (\el) differed significantly among growth forms. Among Raunki\ae r life forms, therophytes (annuals), hemicryptophytes (perennial herbs with buds near the soil surface), and chamaephytes (subshrubs) had greater \els~than phanerophytes (woody plants) and geophytes (perennial herbs with storage organs) (Fig.~\ref{fig:lf-light}; ANOVA - $F_{4, 367}$ = 18.3, $P$ = 1.1 $\times10^{-13}$). Likewise, herbaceous plants (annual, biennial, and perennials) had greater \els~ than shrubs and trees (Fig.~\ref{fig:gf-light}; ANOVA - $F_{4, 367}$ = 10.8, $P$ = 2.6 $\times10^{-8}$)
